Working title: Mirkovi\'c--Vybornov fusion in Beilinson--Drinfeld Grassmannian

### 2020-12-23 examples.tex
sorry i was confusing it

    For the $s$-generalized eigenspace $E_s$, we need $a + sb \neq 0$ to obtain a 2-cycle, which can be taken as 
    \[
    \begin{split}
    & \left\{ 
    (X -sI) \left( e_2 + 2s e_3 + 3s^2 e_4 + \frac{s^3}{a+sb} e_5 + \frac{s^4}{a+sb} e_6 \right), 
    e_2 + 2s e_3 + 3s^2 e_4 + \frac{s^3}{a+sb} e_5 + \frac{s^4}{a+sb} e_6 
    \right\} \\
    = & \left\{ 
    e_1 + s e_2 + s^2 e_3 + s^3 e_4, 
    e_2 + 2s e_3 + 3s^2 e_4 + \frac{s^3}{a+sb} e_5 + \frac{s^4}{a+sb} e_6
    \right\}
    \end{split}
    \]
    The minimal polymonial is $X^3 (X-sI)^2$, which when equated to 0 gives again the equation $cs-bd = 0$.
    Thus the defining equations are 
    \[
    \{ a = 0, cs-bd = 0\}.
    \]
    When we take $s = 0$, we get the equations
    \[
    \{a = 0, bd = 0\}
    \]
    which corresponds to two $\AA^2$'s intersecting along an $\AA^1$. This is indeed an open subset of $\PP^2 \cup_{\PP_1} \PP^2$, as required.
\end{example}

>>>>

    For the $s$-generalized eigenspace $E_s$, we need $a + sb \neq 0$ to obtain a 2-cycle, which can be taken as 
    \[
    \begin{split}
    & \left\{ 
    (X -sI) \left( e_2 + 2s e_3 + 3s^2 e_4 + \frac{s^3}{a+sb} e_5 + \frac{s^4}{a+sb} e_6 \right), \right. \\ 
    & \quad \qquad \qquad  \left. e_2 + 2s e_3 + 3s^2 e_4 + \frac{s^3}{a+sb} e_5 + \frac{s^4}{a+sb} e_6 
    \right\} \\
    = & \left\{ 
    e_1 + s e_2 + s^2 e_3 + s^3 e_4, 
    e_2 + 2s e_3 + 3s^2 e_4 + \frac{s^3}{a+sb} e_5 + \frac{s^4}{a+sb} e_6
    \right\}
    \end{split}
    \]
    The minimal polymonial is $X^3 (X-sI)^2$, which when equated to 0 gives again the equation $cs-bd = 0$.
    Thus the defining equations are 
    \[
    \{ a = 0, cs-bd = 0\}.
    \]
    When we take $s = 0$, we get the equations
    \[
    \{a = 0, bd = 0\}
    \]
    which corresponds to two $\AA^2$'s intersecting along an $\AA^1$. This is indeed an open subset of $\PP^2 \cup_{\PP_1} \PP^2$, as required.
\end{example}

\begin{example}[Example 7 continued\dots]
    The matrix $X$ from the previous example defines (under MVy) the matrix 
    \[
        g = \begin{bmatrix}
            (t-s)t^3 \\
            -bt & (t-s)t \\
            -c & -d & t 
        \end{bmatrix}
    \]
    in $G(\cO)$. 
    Indeed the various blocks of $X$ are in a precise sense the companion matrices of the polynomial entries of $g$
    % First let's double check the inverse map.

    In $\Gr$ the element $g$ defines the lattice 
    \[
        gL_0 = \CC[t]\langle
        (t-s)t^3e_1 - bt e_2 - ce_3 , (t-s)te_2 - de_3, te_3 
        \rangle
    \]
    and computing the matrix of the action of $t$ on the quotient $L_0/L$ in the basis 
    \[
    \{
        [e_1],[te_1],[t^2e_1],[t^3e_1],[e_2],[te_2],[e_3]
    \}    
    \]
    recovers $X$ up to a transpose of course.

    Now let's see what we get when we invert $t$ and $t-s$ respectively. First let's find $ L_s = L\otimes\CC\xt$. 
    \[
        L_s  = \CC[t]\langle(t-s) e_1 - be_2 - ce_3 , (t-s) e_2 , e_3 \rangle = \CC[t]\langle(t-s) e_1 - be_2, (t-s) e_2, e_3 \rangle
    \]
    so 
    \[
    \left[t\big|_{L_0/L_s}\right]_{\{[e_1],[e_2]\}} = \begin{bmatrix}
        s \\
        -b & s 
    \end{bmatrix}
    \]
    which has block type $\mu_2$ and Jordan type $\lambda_2$ as expected.

    Next let's find $L_0 = L\otimes\CC\xt[t-s]$. 
    \[
        L_0 = \CC[t]\langle
        t^3 e_1 - bte_2 - ce_3, te_2 - de_3 , te_3 
        \rangle 
    \]
    so 
    \[
    \left[t\big|_{L_0/L_s}\right]_{\{[e_1],[te_1],[t^2e_1],[e_2],[e_3]\}} = \begin{bmatrix}
        0 \\
        1 & 0 \\
        0 & 1 & 0 \\
          & & & 0 \\
          & & & d & 0  
    \end{bmatrix}
    \]
    which has block type $\mu_1$ and Jordan type $\lambda_1$ as expected. 
\end{example}

### 2020-12-23 main.tex
uncomment examples
sorry i was confusing it

The BD Grassmannian. The convolution Grassmannian. Distinguished orbits, slices therein. Mirkovi\'c--Vybornov \cite{mirkovic2007quiver,mirkovic2019comparison}, Cautis--Kamnitzer \cite{cautis2018categorical}, Anderson--Kogan \cite{anderson2005algebra}.

@@

Let $\Gr$ denote the ordinary affine Grassmannian, $\Grbd$ the Beilinson--Drinfeld affine Grassmannian, and $\Grc$ the convolution affine Grassmannian.

\begin{definition}
    The \new{BD Grassmannian} is the set 
    \begin{equation}
        \label{eq:grbd}
    \begin{aligned}
        \{(V,\sigma) : V &\text{ is a rank } m \text{ vector bundle on }\PP^1 \\ &\text{and } \sigma: V \dashrightarrow \scO_{\PP^1}^m \text{ is a trivialization} \\
        &\text{defined away from finitely many points in }\AA^1\}
    \end{aligned}
    \end{equation}
    More generally, one can define a BD grassmannian over any smooth curve $C$ as the reduced ind-scheme $\Grbd_C$ fibered over a finite symmetric power of $C$ such that the fibre over the point $\vec p$ is a collection of vector bundles over $C$ which are trivial away from $\vec p$ viewed also as a subset of $C$. To \cite{mirkovic2019comparison} the rank $m$ of the trivial fibres $\scO_C^m$ is the of the group $\GL_m\CC$. 
\end{definition}

To quote \cite{baumann2020bases} the BD Grassmannian is a relative version of the affine Grassmannian where the base is the space of effective divisors on a smooth curve. The choice of $\AA^1$ amply satisfies our needs and offers three advantages: there is a natural global coordinate it, every $G$-torsor on it is trivializable, and the monodromy of any local system is trivial. Formally, $\Grbd$ is the functor on the category of commutative $\CC$-algebras that assigns to an algebra $R$ the set of isomorphism classes of triples $(\vec p,V,\sigma)$ where $\vec p\in \AA^n(R)$, $V$ is a $G^\vee$-torsor over $\AA^1_R$ and $\sigma$ is a trivialization of $V$ away from $\vec p$. They denote by $\pi$ the fibration $\Grbd\to\AA^n$ (forgetting $V$ and $\sigma$).

Their simplified description is: it's the set of pairs $(\vec p,[\sigma])$ where $\vec p\in\CC^n$ and $[\sigma]$ is an element of the homogeneous space 
\[
G^\vee(\CC[z,(z-p_1)^{-1},\dots,(z-p_n)^{-1}])/G^\vee(\CC[z])    
\]
Their example, our setting: 
\begin{example}
    When $G = \GL_m\CC$ the datum of $[\sigma]$ is equivalent to the datum of the $\CC[z]$-lattice $\sigma(L_0)$ in $\CC(z)^m$ with $L_0 = \CC[z]^m$ denoting the \new{standard lattice}. Set $f_{\vec p} = (z - p_1)\cdots (z-p_n)$. Then a lattice $L$ is of the form $\sigma(L_0)$ if and only if there exists a positive integer $k$ such that $f^k_{\vec p}(L_0) \subseteq L\subseteq f^{-k}_{\vec p}(L_0)$ and for each $k$ they denote by $\Grbd_k$ the subset of $\Grbd$ consisting of pairs $(\vec p,L)$ such that this sandwhich condition holds. They identify $\CC[z]/(f_{\vec p}^{2k})$ with the vector space of polynomials of degree strictly less than $2kn$, and $L_0/f_{\vec p}^{2k}L_0$ with its $N$th product. Then 
    \[
    \Grbd_k \overset{\text{Zariski closed}}{\subset} \CC^n\times\bigcup_{d=0}^{2knN} G_d(L_0/f_{\vec p}^{2k}L_0)  
    \]
    where $G_d(?)$ denotes the ordinary Grassmann manifold of $d$-planes in the argument.
\end{example}

\begin{definition}
    The \new{deformed convolution Grassmannian} is [not needed?]
    pairs $(\vec p,[\vec\sigma])$ where $\vec p\in\CC^n$ and $\vec\sigma$ is in 
    \[
    G^\vee(\CC[z,(z-p_1)^{-1}]) \times^{G^\vee(\CC[z])}\cdots\times^{G^\vee(\CC[z])} G^\vee(\CC[z,(z-p_n)^{-1}])/G^\vee(\CC[z])
    \] with a map down to $\Grbd$ defined by $(\vec p,[\vec \sigma])\mapsto (\vec p,[\sigma_1\cdots\sigma_n])$. 
\end{definition}
To steal the follow-up example in \cite{baumann2020bases} where the above definition is also copied from\dots 
\begin{example}
    When $G = \GL_m\CC$ this deformation is described by the datum of $\vec p\in\CC^n$ and a sequence $(L_1,\dots,L_n)$ of $\CC[z]$-lattices in $\CC(z)^m$ such that for some $k\in\ZZ$ and for all $j \in \{1\dots n\}$
    \[
    (z-p_j)^kL_{j-1}\subset L_j\subset (z-p_j)^{-k} L_{j-1}    
    \]
    where again $L_0 = \CC[z]^m$ denotes the standard lattice, and now $L_j = (\sigma_1\cdots\sigma_j)(L_0)$. Very nice. Very concrete. They can partition the deformation into \new{cells} by specifying the \new{relative positions} of the pairs $(L_{j-1},L_j)$ in terms of \new{invariant factors}. 
\end{example}
To be continued: \cite{baumann2020bases} go on to describe the fibres of the composition deformation to $\Grbd$ to $\CC^n$ and their description maybe helpful.

@@

@@

Let 
\[
\mu = (\mu^{(1)}, \mu^{(2)}, ..., \mu^{(n)}).
\]
Define $\TT_{\mu_1, \mu_2}$ to be $|\mu| \times |\mu|$ matrices $X$ such that $X$ consists of block matrices where the size of the $i$-th diagonal block is $|\mu^{(i)}| \times |\mu^{(i)}|$, for $1\leq i \leq n$. Each diagonal block is the companion matrix for $t^{\mu_1}(t-s)^{\mu_2}$.

@@

\include{examplesp}

>>>>

The BD Grassmannian. The convolution Grassmannian. Distinguished orbits and slices. Connections to Mirkovi\'c--Vybornov \cite{mirkovic2007quiver,mirkovic2019comparison}, Cautis--Kamnitzer \cite{cautis2018categorical}, Anderson--Kogan \cite{anderson2005algebra}.

@@

Let $\Gr$ denote the ordinary \new{affine Grassmannian} $\Gch(\cK)/\Gch(\cO)$, $\Grbd$ the \new{Beilinson--Drinfeld affine Grassmannian}, and $\Grc$ the \new{convolution affine Grassmannian}.

\begin{definition}
    The \new{BD Grassmannian} is the set 
    \begin{equation}
        \label{eq:grbd}
    \begin{aligned}
        \{(V,\sigma) : V &\text{ is a rank } m \text{ vector bundle on }\PP^1 \\ &\text{and } \sigma: V \dashrightarrow \scO_{\PP^1}^m \text{ is a trivialization} \\
        &\text{defined away from finitely many points in }\AA^1\}
    \end{aligned}
    \end{equation}
    The rank of the $m$ in the definition of $\Grbd$ is the dimension of the maximal torus of $\Gch$. For $\Gch =\GL_m = G$.

    More generally, one can define a BD grassmannian of $\GL_m$ over any smooth curve $C$ as the reduced ind-scheme $\Grbd[n,C]$ fibered over a finite symmetric power of $C$ --- $C^{(n)}$ --- such that the fibre over the point $\vec p = (p_1,\dots,p_n)\in C^n$ is a collection of rank $m$ vector bundles $V$ over $C$ which are trivial away from $\vec p$ viewed also as a subset --- $\{p_1,\dots,p_n\}$ --- of $C$. Trivial means $\scO_C^m\cong V$. 
\end{definition}

To quote \cite{baumann2020bases} the BD Grassmannian is a relative version of the affine Grassmannian where the base is the space of effective divisors on a smooth curve $C$. The choice $C = \AA^1$ ``amply satisfies our needs and offers three advantages: there is a natural global coordinate it, every \anne{$G$-torsor}{i.e. principal $G$ bundle?} on it is trivializable, and the monodromy of any local system is trivial. Formally, $\Grbd$ is the functor on the category of commutative $\CC$-algebras that assigns to an algebra $R$ the set of isomorphism classes of triples $(\vec p,V,\sigma)$ where $\vec p\in \AA^n(R)$, $V$ is a $G^\vee$-torsor over $\AA^1_R$ and $\sigma$ is a trivialization of $V$ away from $\vec p$.''

They denote by $\pi$ the fibration $\Grbd\to\AA^n$ (forgetting $V$ and $\sigma$).
% 
Their simplified description is: it's the set of pairs $(\vec p,[\sigma])$ where $\vec p\in\CC^n$ and $[\sigma]$ is an element of the homogeneous space 
\[
G^\vee(\CC[z,(z-p_1)^{-1},\dots,(z-p_n)^{-1}])/G^\vee(\CC[z])    
\]
Their example, (almost) our setting: 
\begin{example}
    When $G = \GL_m\CC$ the datum of $[\sigma]$ is equivalent to the datum of the $\CC[z]$-lattice $\sigma(L_0)$ in $\CC(z)^m$ with $L_0 = \CC[z]^m$ denoting the \new{standard lattice}. Set $f_{\vec p} = (z - p_1)\cdots (z-p_n)$. Then a lattice $L$ is of the form $\sigma(L_0)$ if and only if there exists a positive integer $k$ such that $f^k_{\vec p}(L_0) \subseteq L\subseteq f^{-k}_{\vec p}(L_0)$ and for each $k$ they denote by $\Grbd_k$ the subset of $\Grbd$ consisting of pairs $(\vec p,L)$ such that this sandwhich condition holds. They identify $\CC[z]/(f_{\vec p}^{2k})$ with the vector space of polynomials of degree strictly less than $2kn$, and $L_0/f_{\vec p}^{2k}L_0$ with its $N$th product. Then 
    \[
    \Grbd_k \overset{\text{Zariski closed}}{\subset} \CC^n\times\bigcup_{d=0}^{2knN} G_d(L_0/f_{\vec p}^{2k}L_0)  
    \]
    where $G_d(?)$ denotes the ordinary Grassmann manifold of $d$-planes in the argument.
\end{example}

Our setting is $G= \GL_m$ and $n = 2$.

\begin{definition}
    The \new{deformed convolution Grassmannian} is [not needed?]
    pairs $(\vec p,[\vec\sigma])$ where $\vec p\in\CC^n$ and $\vec\sigma$ is in 
    \[
    G^\vee(\CC[z,(z-p_1)^{-1}]) \times^{G^\vee(\CC[z])}\cdots\times^{G^\vee(\CC[z])} G^\vee(\CC[z,(z-p_n)^{-1}])/G^\vee(\CC[z])
    \] with a map down to $\Grbd$ defined by $(\vec p,[\vec \sigma])\mapsto (\vec p,[\sigma_1\cdots\sigma_n])$. 
\end{definition}
To steal the follow-up example in \cite{baumann2020bases} where the above definition is also copied from\dots 
\begin{example}
    When $G = \GL_m\CC$ this deformation is described by the datum of $\vec p\in\CC^n$ and a sequence $(L_1,\dots,L_n)$ of $\CC[z]$-lattices in \anne{$\CC(z)^m$}{Why Laurent polynomials for the convolution?} such that for some $k\in\ZZ$ and for all $j \in \{1\dots n\}$
    \[
    (z-p_j)^kL_{j-1}\subset L_j\subset (z-p_j)^{-k} L_{j-1}    
    \]
    where again $L_0 = \CC[z]^m$ denotes the standard lattice, while $L_j = (\sigma_1\cdots\sigma_j)(L_0)$. Very nice. Very concrete. They can partition the deformation into \new{cells} by specifying the \new{relative positions} of the pairs $(L_{j-1},L_j)$ in terms of \new{invariant factors}. 
\end{example}
To be continued: \cite{baumann2020bases} go on to describe the fibres of the \anne{composition}{Not to $\CC^{(n)}$? Or to $\CC$?}
\[
    \Grc \to \Grbd \to \CC^n = \AA^n_\CC 
\] 
their description may be helpful.

@@

Going forward, we'll use $t$ to denote the coordinate on $\AA^1_\CC = \CC$ instead of $z$. Then by $t^\mu \in \Gch(\cK)$ we'll denote the point defined by the coweight $\mu \in \Hom(\CC^\times,T^\vee) = T^\vee(\cK)$ and by $L_\mu$ its image $t^\mu \Gch(\cO)$ in $\Gr$.

\begin{definition}
    Given $\mu_1,\mu_2$ such that $\mu = \mu_1 +\mu_2$ is a partition of $N$ we define $T_{\mu_1,\mu_2}\subset \mat_N$ to be the set of $\mu\times\mu$ block matrices that are zero everywhere except possibly in the last $\min(\mu_i,\mu_j)$ columns of the last row of the $\mu_i\times\mu_j$th block \textit{plus} the block diagonal matrix whose $\mu_i\times\mu_i$ diagonal block is the companion matrix of $t^{\mu_{1,i}}(t-s)^{\mu_{2,i}}$ for each $i\in\{1,2,\dots,m\}$. We call this set \new{name}.
\end{definition}

\begin{remark}
    While we limit ourselves to the case of dominant partitions, the definition above makes sense for arbitrary partitions.
\end{remark}
\begin{remark}
    Speak to whether or not this slice appears in \cite{mirkovic2007quiver}. We don't think it does. But its ``lift'' might. 
\end{remark}

\begin{definition}
    Given $\lambda_1,\lambda_2$ such that $\lambda = \lambda_1 + \lambda_2$ is a partition of $N$ we define $\OO_{\lambda_1,\lambda_2}$ to be the set of $N\times N$ \anne{\new{semi-nilpotent}}{Cute?} matrices $X$ with spectrum in $\{0,s\}$ for some $s\in\CC^\times$ such that 
    $X\big|_{E_0} \in \OO_{\lambda_1}$ and $(X-s)\big|_{E_S}\in\OO_{\lambda_2}$ meaning TODO. 
    % restricted to its generalized $0$ eigenspace $X$ is nilpotent of Jordan type $\lambda_1$ and restricted to its generalized $s$ eigenspace $X - s$...
\end{definition}

Correspondingly we have 
\begin{itemize}
    \item $W_{\mu_1,\mu_2} = \Gch_1 [t^{-1},(t-s)^{-1}]t^{\mu_1} (t-s)^{\mu_2} \Gch(\cO)$ as a subset of $\Grbd[2]$ where recall the subscript 2 records the fact that we are fixing two points $0,s\in\CC$
    \item Does $\Grbd[2]^{\lambda_1,\lambda_2}$ admit such an orbit description? As a fibration it is pairs $(V,\sigma)$ such that is trivialized away from $(0,s)$ by $\sigma$ and $\sigma$ has type $(\lambda_1,\lambda_2)$ --- the data of the trivialization is equivalent to the data of pairs of lattices $(L_1,L_2)$ such that $L_i \in \Gr^{\lambda_i}$ away from $0$ and $L_1 = L_2\in\Gr^{\lambda}$ at 0?  
    \item In the deformed convolution Gr we have the subset $\Grc[2]^{\lambda_1,\lambda_2}$ of pairs $((0,s),[\sigma_1,\sigma_2])$ (really want arbitrary $(p_1,p_2)$ in place of $(0,s)$ which is the specialization that we make when we work in $\Grbd[2]$ sort of?) with $\sigma_i\in\Gch(\CC[t])(t-p_i)^{\lambda_i}\Gch(\CC[z])$ but this is not important? 
\end{itemize}

\begin{question}
    Can we describe $W_{\mu_1,\mu_2}\cap \overline{\Gr^{\lambda_1,\lambda_2}}$ as the set of lattices $L$ such that 
    \[
        t^{\lambda_{1,1}}(t-s)^{\lambda_{2,1}}L_0 \subset L \subset t^{-\lambda_{1,1}}(t-s)^{-\lambda_{2,1}}L_0
    \]
    and $\lim_{s\to 0} \rho^\vee (s)\cdot L = L_\mu$
\end{question}

@@

Let 
\[
\mu = (\mu^{(1)}, \mu^{(2)}, ..., \mu^{(n)}).
\]
Define $\TT_{\mu_1, \mu_2}$ to be $|\mu| \times |\mu|$ matrices $X$ such that $X$ consists of block matrices where the size of the $i$-th diagonal block is $|\mu^{(i)}| \times |\mu^{(i)}|$, for $1\leq i \leq n$. Each diagonal block is the companion matrix for $t^{\mu_1}(t-s)^{\mu_2}$. Each off-diagonal block is zero everywhere except possibly in the last $\min(\mu_i,\mu_j)$ columns of the last row.

@@

\include{examples}
